---
title: "Harpeth Conservancy's E. coli Forecast"
---

```{r warning=FALSE, setup, include=FALSE}

library(rmarkdown)
library(reticulate)
library(tidyverse)
library(leaflet)
library(htmlwidgets)
library(widgetframe)
library(leaflet.providers)
library(sf)
library(leaflet.extras)

Sys.which("python")

use_python("C:/Users/rwjac/miniconda/python.exe")


```

```{python}
!pip install netCDF4
!pip install wget
!pip install pyproj
```

```{python}
import json
import pandas as pd
import requests
import numpy as np
from os import replace
import datetime
from datetime import datetime, timedelta
import math
import netCDF4 as nc
import wget
import pytz
```


```{python}
time = datetime.now(pytz.timezone('US/Central'))

month = time.month
year = time.year
day = time.day

mm = time.strftime('%m')
dd = time.strftime('%d')
stime = time.strftime('%Y%m%d')

weekago_month = (time - timedelta(days=7)).month
weekago_year = (time - timedelta(days=7)).year
weekago_day = (time - timedelta(days=7)).day

day_of_year = time.timetuple().tm_yday
B = 0.01721420632
C = 81.75
sindoy = math.sin(B * (day_of_year) - C)
sindoy_list = [sindoy]*8
```

Pulling Temperature forecast data for each day from NWS's API.

```{python}

def get_temp_dict(url):
  request = (requests.get(url).text)
  temp_dict = json.loads(request)
  temp_dict = temp_dict['properties']['periods'][0]
  return temp_dict

def get_temp_df(temp_dict):
  temp_df = pd.DataFrame.from_dict(temp_dict)
  temp_df = temp_df[['temperature', 'probabilityOfPrecipitation','windSpeed', 'windDirection']]
  temp_df.columns = ['temperature', 'precipitation', 'windSpeed', 'windDirection']
  return temp_df

def get_temp_forecast(temp_df):
  temp_df['temperature'] = temp_df['temperature'].astype(str).astype(int)
  temp_forecast = temp_df['temperature'][1]
  return temp_forecast

def pull_temp(url):
  temp_dict = get_temp_dict(url)
  temp_df = get_temp_df(temp_dict)
  temp_forecast = get_temp_forecast(temp_df)
  return temp_forecast

```

```{python}

OHX_values = ['45,52', '39,54', '47,54', '48,45', '54,51', '48,55', '52,55', '46,50']

temp_urls = []

for x in OHX_values:
  temp_urls.append("https://api.weather.gov/gridpoints/OHX/{}/forecast".format(x))
  
```



```{python}

temp_list = []

for x in temp_urls:
  temp_forecast = pull_temp(x)
  temp_list.append(temp_forecast)

print(temp_list)

```

Pulling Flow Data from USGS

```{python}

flow_sites = ['03431700', '03433500', '03434500', '03432350', '03431060', '03434500', '03431700', '03431060']

flow_urls = []

for x in flow_sites:
  flow_urls.append("https://waterservices.usgs.gov/nwis/iv/?format=json&sites={}&startDT={}-{}-{}T04:00-0500&endDT={}-{}-{}T05:00-0500&parameterCd=00060&siteStatus=all".format(x, year, month, day, year, month, day).format(x))

```

```{python}

def pull_floworgage(url):
  data = json.loads(requests.get(url).text)
  data = data['value']['timeSeries'][0]['values'][0]['value']
  df = pd.DataFrame.from_dict(data)
  df['value'] = df['value'].astype(str).astype(float)
  floworgage = df['value'].mean()
  return floworgage

```


```{python}

flow_list = []

for x in flow_urls:
  flow = pull_floworgage(x)
  flow_list.append(flow)

print(flow_list)

```
Pulling precipitation data

The precipitation data are quality-controlled, multi-sensor (radar and rain gauge) precipitation estimates obtained from National Weather Service (NWS) River Forecast Centers (RFCs) and mosaicked by National Centers for Environmental Prediction (NCEP). 

```{python}
def get_dataset(url):
  retrieved_file = wget.download(url)
  dataset = nc.Dataset(retrieved_file)
  return dataset

def get_time(url):
  dataset = get_dataset(url)
  time_array = dataset['time'][:].toflex()
  time = time_array['_data'][()]
  datetime_obj = datetime.fromtimestamp(time)
  return datetime_obj.isoformat()

def create_prcp_array(dataset):
  prcp_array = dataset['observation'][473:487, 789:796]
  prcp_array = prcp_array.toflex()
  return prcp_array

def get_prcp(array):
  prcp = []
  xcrsfordf = []
  ycrsfordf = []
  for i in range (0, 14):
    for n in range (0, 7):
     prcp.append(array[i]['_data'][n])
     xcrsfordf.append(789 + n)
     ycrsfordf.append(473 + i)
  return prcp, xcrsfordf, ycrsfordf

def create_df(prcp, xcrsfordf, ycrsfordf):
  df = pd.DataFrame()
  df['Precipitation'] = prcp
  df['X-coordinates'] = xcrsfordf
  df['Y-coordinates'] = ycrsfordf
  return df

def find_prcp(dataframe, xcrs, ycrs):
  prcp_list = []
  for (a,b) in zip(xcrs, ycrs):
    value = dataframe.loc[dataframe['X-coordinates'].eq(a) & dataframe['Y-coordinates'].eq(b), 'Precipitation']
    prcp_value = value.values[0]
    prcp_list.append(prcp_value)
  return prcp_list

def pull_prcp(url, xcrs, ycrs):
  dataset = get_dataset(url)
  prcparray = create_prcp_array(dataset)
  prcp_info = get_prcp(prcparray)
  df = create_df(prcp_info[0], prcp_info[1], prcp_info[2])
  prcp_list = find_prcp(df, xcrs, ycrs)
  return prcp_list

```

```{python}
prcp_urls = ['https://water.weather.gov/precip/downloads/{}/{}/{}/nws_precip_1day_{}_conus.nc'.format(year, mm, dd, stime),
        'https://water.weather.gov/precip/downloads/{}/{}/{}/nws_precip_last2days_{}_conus.nc'.format(year, mm, dd, stime),
        'https://water.weather.gov/precip/downloads/{}/{}/{}/nws_precip_last3days_{}_conus.nc'.format(year, mm, dd, stime),
        'https://water.weather.gov/precip/downloads/{}/{}/{}/nws_precip_last5days_{}_conus.nc'.format(year, mm, dd, stime),
        'https://water.weather.gov/precip/downloads/{}/{}/{}/nws_precip_last7days_{}_conus.nc'.format(year, mm, dd, stime)]

xcrs = [792, 789, 792, 795, 795, 792, 792, 793]
ycrs = [484, 486, 479, 486, 476, 482, 478, 473]

prcp_lists = []

for x in prcp_urls:
   prcp_list = pull_prcp(x, xcrs, ycrs)
   prcp_lists.append(prcp_list)
   
onedayprcp_list = prcp_lists[0]
twodayprcp_list = prcp_lists[1]
threedayprcp_list = prcp_lists[2]
fivedayprcp_list = prcp_lists[3]
weekprcp_list = prcp_lists[4]
#timestamp = get_time(urls[0])
```

Pulling Gage Height data from USGS

```{python}

gage_sites = ['03431700', '03433500', '03434500', '03432350', '03431060', '03434500', '03431700', '03431060']

gage_urls = []

for x in gage_sites:
  gage_urls.append("https://waterservices.usgs.gov/nwis/iv/?format=json&sites={}&startDT={}-{}-{}T04:00-0500&endDT={}-{}-{}T05:00-0500&parameterCd=00065&siteStatus=all".format(x, year, month, day, year, month, day).format(x))

```


```{python}

gage_list = []

for x in gage_urls:
  gage = pull_floworgage(x)
  gage_list.append(gage)

print(gage_list)

```

Creating a DataFrame of Forecast Values

```{python}

column_names = ['temp', 'onedayprcp', 'twodayprcp', 'threedayprcp', 'fivedayprcp', 'weekprcp', 'flow', 'gage', 'sin(dayofyear)']

```


```{python}

index_names = ['Hwy100', 'Hwy70', 'JBlvd', 'LPike', 'MCreek', 'MBridge', 'RCreek', 'WPark']

```


```{python}

forecast_df = pd.DataFrame(list(zip(temp_list, onedayprcp_list, twodayprcp_list, threedayprcp_list, fivedayprcp_list, weekprcp_list, flow_list, gage_list, sindoy_list)), index = index_names, columns = column_names)

```

```{r warning=FALSE, models}

# use Python data and convert into R dataframe
forecast_df<-py$forecast_df

#Add location description for use in map
forecast_df$location <- c("Highway 100 Boat Launch", "Hwy 70 Boat Launch", "Jackson Blvd", "Lewisburg Pike", "Mill Creek Greenway", "Moran Road Bridge", "Richland Creek Greenway", "Whitsett Park")



# add latitude and longitude for each sampling location
# Sampling location order: Hwy100, Hwy70, JBlvd, LPike, MCreek, MBridge, RCreek, WPark

Lat<-c('36.054361', '36.123611',  '36.112694', '35.909389',  '36.017250', '36.017167', '36.132250', '36.118361')

Long<-c('-86.928806', '-87.098944', '-86.862500','-86.855806',   '-86.686250', '-86.900139', '-86.848667', '-86.724333')

forecast_df$Latitude<-Lat
forecast_df$Longitude<-Long

#Hwy100 Model - Ecoli = 207.2 + 0.06577*(PROD(Flow,prev_two_rainfall)) - 61.21*(PROD(sinDOY,GH))

Hwy100_pred <- 207.2+0.06577*(forecast_df["Hwy100","flow"]*forecast_df["Hwy100","twodayprcp"])-61.21*(forecast_df["Hwy100","sin(dayofyear)"]*forecast_df["Hwy100","gage"])

#Hwy70 Model - Ecoli = -592.2 + 0.1694*(PROD(Flow,prev_two_rainfall)) + 1.973e+04*(INVERSE(SUM(sinDOY,TAVG),6.2684246))
#THIS MODEL IS CURRENTLY WRONG!!
Hwy70_pred <- -592.2 + 0.1694*(forecast_df["Hwy70","flow"]*forecast_df["Hwy70","twodayprcp"]) + 1.973e+04*(1/(forecast_df["Hwy70","sin(dayofyear)"] + forecast_df["Hwy70","temp"]))+500

#JBlvd Model - Ecoli = -312.7 + 136*(SQUAREROOT(rainfall)) + 2.146e+04*(INVERSE(TMAX,8.35)) + 144.8*(INVERSE(PROD(sinDOY,TAVG),-1.2868741))

JBlvd_pred <- -312.7+ 136 * sqrt(forecast_df["JBlvd","onedayprcp"]) + forecast_df["JBlvd","temp"]+2000

#LPike Model - Ecoli = 294.9 - 115.9*(sinDOY) + 3.068e-05*(SQUARE(PROD(Flow,prev_two_rainfall)))

LPike_pred <- 294.9 - 115.9*forecast_df["LPike","sin(dayofyear)"] + 3.068e-05*((forecast_df["LPike","flow"]*forecast_df["LPike","twodayprcp"])^2)

#MCreek Model - Ecoli = -15.33 + 5.972*(SUM(Flow,prev_two_rainfall))

MCreek_pred <- -15.33 + 5.972*(forecast_df["MCreek","flow"]+forecast_df["MCreek","twodayprcp"])

#MBridge Model - Ecoli = 170.9 + 0.5829*(PROD(Flow,GH)) - 97.66*(SQUAREROOT(sinDOY))

MBridge_pred <- 170.9 + 0.5829*(forecast_df["MBridge","flow"]*forecast_df["MBridge","gage"]) - 97.66*sqrt(forecast_df["MBridge","sin(dayofyear)"])

#RCreek Model - Ecoli = -191.8 + 851.4*(LOG10(SUM(rainfall,Flow)))

RCreek_pred <- -191.8 + 851.4* log10(forecast_df["RCreek","onedayprcp"]+forecast_df["RCreek","flow"])

#WPark Model - Ecoli = 7.295 + 6.334*(SUM(Flow,prev_two_rainfall))

WPark_pred <- 7.295 + 6.334*(forecast_df["WPark","flow"]+forecast_df["WPark","twodayprcp"])

forecast_df$Ecoli_pred <- round(c(Hwy100_pred, Hwy70_pred, JBlvd_pred, LPike_pred, MCreek_pred, MBridge_pred, RCreek_pred, WPark_pred))


```

```{r warning=FALSE, status}

forecast_df2 = forecast_df

forecast_df$status[forecast_df2$Ecoli_pred < 150] <- "Safe"
forecast_df$status[forecast_df2$Ecoli_pred > 150 & forecast_df2$Ecoli_pred < 250] <- "Caution"
forecast_df$status[forecast_df2$Ecoli_pred > 250 & forecast_df2$Ecoli_pred < 500] <- "Warning"
forecast_df$status[forecast_df2$Ecoli_pred > 500] <- "Danger"

forecast_df$status[is.na(forecast_df$Ecoli_pred)] = NA


```



```{r warning=FALSE, map}

forecast_df %<>% 
  st_as_sf(coords = c("Longitude", "Latitude")) %>% 
  st_sf(crs = 84)

forecast_df<-forecast_df %>% 
  mutate(popup = paste("<b>Location: </b>", forecast_df$location, 
                       "<br/>",
                       "<b>E. coli:</b>", forecast_df$Ecoli_pred,
                       "<br/>",
                       "<b>Status:</b>", forecast_df$status,
                       "<br/>"))



cfac<-colorFactor(c("green", "yellow", "orange", "red"), levels=c("Safe","Caution","Warning","Danger"),na.color = "grey", ordered=F)


p<-leaflet() %>% 
  addTiles() %>% 
  addProviderTiles(providers$OpenStreetMap, group = 'Open SM')  %>%
  addProviderTiles(providers$Stamen.Toner, group = 'Toner')  %>%
  addProviderTiles(providers$Esri.NatGeoWorldMap, group = 'NG World') %>%
  setView(lng = -86.791425, lat = 36.08, zoom = 10) %>% 
  addCircleMarkers(data = forecast_df ,popup = ~popup, color=~cfac(forecast_df$status),fill=T, stroke=T, fillOpacity = 1.0) %>%
  addLegend("topright", 
            colors = c("green",  "yellow", "orange", "red", "gray"),
            labels = c("Safe","Caution","Warning", "Danger", "Data Unavailable"),
            title = 'Public Safety Status',
            opacity = 1) %>%
  leaflet.extras::addResetMapButton() %>% 
  addScaleBar(
    position = "bottomright",
    options = scaleBarOptions(imperial = TRUE)
    ) %>%
  addLayersControl(
    position = "bottomleft",
    baseGroups = c("Open SM","Toner","NG World"),
    # Choose to permanently display or collapse layers control switch
    options = layersControlOptions(collapsed = FALSE)
    )  




p

```


```{r warning=FALSE, mapsave, include=FALSE}
#widget_file_size <- function(p) {
#  d <- tempdir()
#  withr::with_dir(d, htmlwidgets::saveWidget(p, "map.html"))
#  f <- file.path(d, "map.html")
#  mb <- round(file.info(f)$size / 1e6, 3)
#  message("File is: ", mb," MB")
#}
#widget_file_size(p)

#widget_file_size(partial_bundle(p))


#saveWidget(here::here('/files/html/posts', 'predoc_map.html'))

saveWidget(frameableWidget(p), "map.html", selfcontained = T, libdir = "lib")

```

